---
title: "Cecilia's Portfolio 2025"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme: "journal"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggplot2)
library(rjson) 
library(plotly)
source("compmus.R")
```

```{r}
library(tidymodels)
library(ggdendro)
library(heatmaply)
```
## Computational Musicology Portfolio 2025

***

Welcome :)


### Comparing dendrograms of class corpus week 12

#### Dendrogram with all features

*** 

So for this assignment I decided to compare two dendrograms of the class corpus. The first one is with all the features (arousal, danceability, instrumentalness, tempo and valence), as for the second only has arousal and valence. My prediction is that if you get rid of some of the features, the song will become less different. Lets find out!

```{r}
compmus2025 <- read.csv("compmus2025 (2).csv")

```

```{r}
source("compmus.R")

get_conf_mat <- function(fit) {
  outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
}  
```

```{r}
cluster_juice <-
  recipe(
    filename ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(compmus2025) |>
  juice() |>
  column_to_rownames("filename")

compmus_dist <- dist(cluster_juice, method = "euclidean")

compmus_dist |> 
  hclust(method = "average") |>
  dendro_data() |> 
  ggdendrogram() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 5))
```

#### Dendrogram with only arousal and valence

```{r}
cluster_juice <-
  recipe(
    filename ~
      arousal +
      valence,
    data = compmus2025
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(compmus2025) |>
  juice() |>
  column_to_rownames("filename")

compmus_dist <- dist(cluster_juice, method = "euclidean")

compmus_dist |> 
  hclust(method = "average") |>
  dendro_data() |> 
  ggdendrogram() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 5))
```

***

As you can see, the dendrograms look different. In the first one, the y-axis counts till 4 (and even over 4), as in the second one it doesn't get much higher than two. The tracks are still all very different compared to eachother, but my hypothesis was right that they are a bit more similar compared to if you use all the features

### AI or non AI?

#### Prediction

***

Here we're going to look at the prediction. Does the computer think the tracks are AI or non AI? 

```{r}
compmus2025_filtered <- 
  compmus2025 |> filter(!is.na(ai)) |> 
  mutate(ai = factor(if_else(ai, "AI", "Non-AI")))

classification_recipe <-
  recipe(
    ai ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025_filtered
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors())     

compmus_cv <- compmus2025_filtered |> vfold_cv(5)

knn_model <-
  nearest_neighbor(neighbors = 1) |>
  set_mode("classification") |> 
  set_engine("kknn")
classification_knn <- 
  workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(knn_model) |> 
  fit_resamples(compmus_cv, control = control_resamples(save_pred = TRUE))

classification_knn |> get_conf_mat()
```
```{r}
classification_knn |> get_conf_mat() |> autoplot(type = "mosaic")
```

***

I'm not really sure how I'm supposed to read this...

```{r}
classification_knn |> get_conf_mat() |> autoplot(type = "heatmap")
```

#### Forest-model

```{r}
forest_model <-
  rand_forest() |>
  set_mode("classification") |> 
  set_engine("ranger", importance = "impurity")
indie_forest <- 
  workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(forest_model) |> 
  fit_resamples(
    compmus_cv, 
    control = control_resamples(save_pred = TRUE)
  )

indie_forest |> get_pr()

workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(forest_model) |> 
  fit(compmus2025_filtered) |> 
  pluck("fit", "fit", "fit") |>
  ranger::importance() |> 
  enframe() |> 
  mutate(name = fct_reorder(name, value)) |> 
  ggplot(aes(name, value)) + 
  geom_col() + 
  coord_flip() +
  theme_classic() +
  labs(x = NULL, y = "Importance")
```
```{r}
compmus2025_filtered |>
  ggplot(aes(x = valence, y = arousal, colour = ai, size = tempo)) +
  geom_point(alpha = 0.8) +
  scale_color_viridis_d() +
  labs(
    x = "Valence",
    y = "Arousal",
    size = "Tempo",
    colour = "AI"
  )
```



### Tempograms week 11

#### Tempogram Track 1

```{r}
"features/cecilia-b-1.json" |>
    compmus_tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) |>
    ggplot(aes(x = time, y = bpm, fill = power)) +
    geom_raster() +
    scale_fill_viridis_c(guide = "none") +
    labs(x = "Time (s)", y = "Tempo (BPM)") +
    theme_light()
```

***
This graph of my first track shows how the speed of the music changes over time. Most of the time, the tempo stays steady at around 200 BPM and 400 BPM, shown by the bright, straight lines. At the beginning and end, the lines look more wavy, which means the tempo is changing or the rhythm is less steady in those parts. This happened in this song because it has a small intro that gradually settles into a steady beat and a tempo shift near the end.

#### Tempogram Track 2

```{r}
"features/cecilia-b-2.json" |>
    compmus_tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) |>
    ggplot(aes(x = time, y = bpm, fill = power)) +
    geom_raster() +
    scale_fill_viridis_c(guide = "none") +
    labs(x = "Time (s)", y = "Tempo (BPM)") +
    theme_classic()
```

***
This graph of my second track shows how the tempo changes over time. The bright lines staying in the same place mean the tempo is very steady, around 200 BPM, 400 BPM, and 600 BPM. This suggests the music has a strong, consistent beat without big changes, which is common in electronic or dance music, and my goal was to make a 'dance music' track. The beat stays the same throughout the track.

### Novelty Function: Energy

#### Energy Novelty Track 1

```{r}
"features/cecilia-b-1.json" |>
    compmus_energy_novelty() |> 
    ggplot(aes(t, novelty)) +
    geom_line() +
    theme_minimal() +
    labs(x = "Time (s)", y = "Energy Novelty")
```

***
As you can see, this track doesn't have a lot of energy whatsoever. This track contains chill, relaxing ukulele music so 'low energy' was also my goal. It is kind of boring and as you can see, there's no energy to be found!

#### Energy Novelty Track 2

```{r}
"features/cecilia-b-2.json" |>
    compmus_energy_novelty() |> 
    ggplot(aes(t, novelty)) +
    geom_line() +
    theme_minimal() +
    labs(x = "Time (s)", y = "Energy Novelty")
```

***
My second track is a bit more energetic compared to my first one, but less than I imagined. You can clearly see the 'beat-drop' 20 seconds into the track, but afterwards there is also hardly any energy in this track

### Novelty function: spectral

#### Spectral Novelty Track 1

```{r}
"features/cecilia-b-1.json" |>
    compmus_spectral_novelty() |> 
    ggplot(aes(t, novelty)) +
    geom_line() +
    theme_minimal() +
    labs(x = "Time (s)", y = "Spectral Novelty")
```

***
If you look at this graph of my first track, you can see that it's almost completely the same during the whole song. 

#### Spectral Novelty Track 2

```{r}
"features/cecilia-b-2.json" |>
    compmus_spectral_novelty() |> 
    ggplot(aes(t, novelty)) +
    geom_line() +
    theme_minimal() +
    labs(x = "Time (s)", y = "Spectral Novelty")
```

***
Here you can see the beat drop that happens around 20 seconds into the track. There is an enormous shift in the graph, and after 50 seconds or so it calms down again.

### Chordogram week 10

#### Chordogram Track 1

```{r}
#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )
```


```{r}
"features/cecilia-b-1.json" |> 
  compmus_chroma(norm = "manhattan") |> 
  compmus_match_pitch_templates(
    chord_templates,         
    norm = "manhattan",        
    distance = "manhattan"   
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(option = "plasma", guide = "none") +               
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()  
```

***

In this chordogram, you can see that specific chords were played throughout the entire song. F- and C major are the most obvious ones. This track also contains a ukulele playing the chords, so the chordogram looks pretty logic to me.

#### Chordogram Track 2

```{r}
#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )
```


```{r}
"features/cecilia-b-2.json" |> 
  compmus_chroma(norm = "manhattan") |> 
  compmus_match_pitch_templates(
    chord_templates,         
    norm = "manhattan",        
    distance = "manhattan"   
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(option = "plasma", guide = "none") +               
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()  
```

***

So if you compare this chordogram to the first one, you can see that there are less chords. The colors aren't as clear and not as long as the blocks in track 1. When listening to the track this makes sense, because I can't really hear many chords whatsoever.

Oh and if you read this and know how I can put my texts next to the graphs in stead of under them: please send me a message or add it to the comments! Very much appreciated :)

### Chromagram Analysis

#### Chromagram Track 1

```{r}
"features/cecilia-b-1.json" |>                           
          compmus_chroma(norm = "identity") |>                 
          ggplot(aes(x = time, y = pc, fill = value)) + 
          geom_raster() +
          scale_y_continuous(
                  breaks = 0:11,
                  minor_breaks = NULL,
                  labels = c(
                          "C", "C#|Db", "D", "D#|Eb",
                          "E", "F", "F#|Gb", "G",
                          "G#|Ab", "A", "A#|Bb", "B"
                      )
              ) +
          scale_fill_viridis_c(guide = "none") +               
          labs(x = "Time (s)", y = NULL, fill = NULL) +
          theme_light() 
```

***

Figure: Chromagram of Cecilia B-1

This chromagram visualizes the pitch class distribution over time for my first track. The x-axis represents time in seconds, while the y-axis corresponds to the 12 pitch classes (C to B), including enharmonic equivalents (e.g., C#|Db). The color intensity indicates the strength of each pitch class at a given moment, with brighter areas representing stronger presence. This visualization helps analyze harmonic content, tonality, and key changes throughout the track.

(I still want to analyze it further but ran out of time for now)

#### Chromagram Track 2

```{r}
"features/cecilia-b-2.json" |>                           
          compmus_chroma(norm = "identity") |>                 
          ggplot(aes(x = time, y = pc, fill = value)) + 
          geom_raster() +
          scale_y_continuous(
                  breaks = 0:11,
                  minor_breaks = NULL,
                  labels = c(
                          "C", "C#|Db", "D", "D#|Eb",
                          "E", "F", "F#|Gb", "G",
                          "G#|Ab", "A", "A#|Bb", "B"
                      )
              ) +
          scale_fill_viridis_c(guide = "none") +               
          labs(x = "Time (s)", y = NULL, fill = NULL) +
          theme_light() 
```

***

Figure: Chromagram of Cecilia B-2

This chromagram represents the pitch class distribution over time for my second track. The x-axis shows time in seconds, while the y-axis corresponds to the 12 pitch classes (C to B), including enharmonic equivalents. 

(I still want to analyze it further, and compare both chromagrams but ran out of time for now)

### Cepstrogram Analysis

#### Cepstrogram Track 1

```{r}
"features/cecilia-b-1.json" |>                           
    compmus_mfccs(norm = "identity") |>                  
    ggplot(aes(x = time, y = mfcc, fill = value)) + 
    geom_raster() +
    scale_y_continuous(
        breaks = 0:12,
        minor_breaks = NULL,
    ) +
    scale_fill_viridis_c(option = "plasma", guide = "none") +
    labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
    theme_light() 
```

***

Figure: Cepstrogram of Cecilia B-1

This cepstrogram visualizes the evolution of Mel-Frequency Cepstral Coefficients (MFCCs) over time for my first track. The x-axis represents time in seconds, while the y-axis indicates the MFCC coefficient number, which captures different aspects of the spectral envelope. The color intensity (using the plasma color scale) shows the strength of each coefficient at a given moment, with brighter colors representing higher magnitudes. MFCCs are commonly used for timbre analysis, making this plot useful for understanding tonal quality, texture, and distinguishing different sound characteristics within the track.

#### Cepstrogram Track 2

```{r}
"features/cecilia-b-2.json" |>                           
    compmus_mfccs(norm = "identity") |>                  
    ggplot(aes(x = time, y = mfcc, fill = value)) + 
    geom_raster() +
    scale_y_continuous(
        breaks = 0:12,
        minor_breaks = NULL,
    ) +
    scale_fill_viridis_c(option = "plasma", guide = "none") +
    labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
    theme_light() 
```

***

Figure: Cepstrogram of Cecilia B-2

This cepstrogram displays the progression of Mel-Frequency Cepstral Coefficients (MFCCs) over time for my second track.

(I still want to analyze it further, and compare both cepstrograms but ran out of time for now)

### Self-similarity Matrices Track 1

#### Chroma-based self-similarity matrix track 1

```{r}
"features/cecilia-b-1.json" |>                           
     compmus_chroma(norm = "manhattan") |>                 
     compmus_self_similarity(
         feature = pc,
         distance = "manhattan"                             
     ) |>   
     ggplot(aes(x = xtime, y = ytime, fill = d)) + 
     geom_raster() +
     scale_fill_gradient(low = "purple", high = "pink") +               
     labs(x = "Time (s)", y = NULL, fill = NULL) +
     theme_light() 
```

***

This visualization helps analyze harmonic consistency and motif recurrence within the track.

#### Timbre-based self-similarity matrix track 1

```{r}
"features/cecilia-b-1.json" |>                           
     compmus_mfccs(norm = "identity") |>                  
     compmus_self_similarity(
         feature = mfcc,
         distance = "euclidean"                             
     ) |>   
     ggplot(aes(x = xtime, y = ytime, fill = d)) + 
     geom_raster() +
     scale_fill_gradient(low = "#FF69B4", high = "#FFD700") +               
     labs(x = "Time (s)", y = NULL, fill = NULL) +
     theme_light()
```

*** 

This timbre-based self-similarity matrix (SSM) uses Mel-Frequency Cepstral Coefficients (MFCCs) to measure the similarity in timbre across different points in my first track. The x- and y-axes represent time, while the color gradient illustrates the degree of similarity—brighter regions indicate similar timbral qualities between different sections. Unlike the chroma-based matrix, this SSM captures changes in texture, instrumentation, and dynamics, making it useful for identifying shifts in sound color rather than harmonic structure.

### Self-similarity Matrices Track 2

#### Chroma-based self-similarity matrix track 2

```{r}
"features/cecilia-b-2.json" |>                           
     compmus_chroma(norm = "manhattan") |>                 
     compmus_self_similarity(
         feature = pc,
         distance = "manhattan"                             
     ) |>   
     ggplot(aes(x = xtime, y = ytime, fill = d)) + 
     geom_raster() +
     scale_fill_gradient(low = "purple", high = "pink") +               
     labs(x = "Time (s)", y = NULL, fill = NULL) +
     theme_light() 
```

#### Timbre-based self-similarity matrix track 2

```{r}
"features/cecilia-b-2.json" |>                           
     compmus_mfccs(norm = "identity") |>                  
     compmus_self_similarity(
         feature = mfcc,
         distance = "euclidean"                             
     ) |>   
     ggplot(aes(x = xtime, y = ytime, fill = d)) + 
     geom_raster() +
     scale_fill_gradient(low = "#FF69B4", high = "#FFD700") +               
     labs(x = "Time (s)", y = NULL, fill = NULL) +
     theme_light()
```

### Introduction

```{r}
```

***

Hello everyone, and welcome to my Portfolio page! This is the first time coding for me, and I've been sick all week and missing both the lecture and the workgroup, so I think I know what I'm doing, but not entirely sure... Feedback is very welcome :) 

For the homework assignment of last week I had to create two tracks. I made them through a generative AI program called 'Stable Audio'. I wanted to make two completely different tracks.

Track 1: for this track I wanted to go for a relaxing vibe. I used the prompts ukulele, beach vibes, relax, 80 BPM, high-hat, danceable, cool, build-up and ukulele solo. The result was kind of what I expected: relaxing, summer vibes with a piña colada, lying down on the beach on holiday. The tempo is kind of slow, there are many consonant harmonies, gentle strumming patterns and it's quite minimalistic.

Track 2: this track is a bit more outgoing than the first one. I wanted to make a tune that you'll hear in the club and dance to. Therefor I used the prompts techno club, classical violin, piano chords, cool, groovy, 100 BPM, danceable, high-hat, build-up and heavy beatdrop. As you can see I also used classical music prompts. I like to experiment with 'modern' genres and classical music because I think they make an excellent combination. On this one I think the danceability, energy and instrumentalness pop out the most.

### Visualization week 9

```{r}
 library(ggplot2)
library(plotly)

data <- read.csv("aisc2024.csv")

tempo_danceability_gg <- ggplot(data, aes(x = tempo, y = danceability, size = instrumentalness, color = instrumentalness)) +
    geom_point(alpha = 0.6) + 
    labs(title = "Tempo vs Danceability", 
         x = "Tempo (BPM)", y = "Danceability", size = "Instrumentalness", color = "Instrumentalness") +
    scale_color_viridis_c() +  
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))

ggplotly(tempo_danceability_gg)

```


***

For this week, we were supposed to use the "compmus2025.csv" file to explore them, but I didn't manage to get the file working: it kept saying that the whole document was one column, so I couldn't examine all the separate musical characteristics. 

Ashley said it was also okay to use the "aisc2024.csv" file this week, so that's what I've done. I've looked at all the different features and found especially the tempo and danceability interesting to look at. These are also features where I focused on in my second track. As you can see in the interactive(!) graph, you have tracks with 60 BPM with a high danceability, but also tracks with a BPM of over 150 and a very low danceability. I presumed beforehand that the tracks with a low BPM were not as danceable as the ones with a higher one, so I found this very interesting. 


### Discussion

***

As I said in the visualization, I couldn't use the new file with the Essentia features, so I couldn't see if Essentia characterized my tracks correctly. I hope that it will work next week and I can analyze my own tracks! :)


